---
title: "Salaries of data related jobs from 2020 - 2024"
author: "Morgan"
output: html_document
date: "2025-02-04"
---

## Import Package
```{r, message=F, warning=F}
library(tidyverse)
library(dplyr)
library(psych)
library(stringr)
library(ggplot2)
```

---

## Read csv
```{r}
# read
df <- read.csv("data/DataScience_salaries_2024.csv")
#head(df, 5)

# data type
str(df)

# simple stats
describe(df)

# na value
colSums(is.na(df))  #no na values in each columns
```
---

## Data Description
### Categorical
- **work_year**: the year of the data (2020 - 2024)
- **experience_level**: EN(entry), MI(mid), SE(senior), EX(executive)
- **employment_type**: PT(part-time), FT(full-time), FL(freelance), CT(contract)
- **job_title**: the title of the job
- **salary_currency**: the currency in which the salary is denoted
- **employee_residence**: the residence location of employee
- **company_location**: the location of the company
- **company_size**: S(small), M(medium), L(large)

### Numerical
- **salary**: the salary of the employee (14000 - 30400000)
- **salary_in_usd**: salary converted to USD (15000 - 800000)
- **remote_ratio**: the ratio of remote work allowed (0 - 100)

---

## Modification
### M1: map all abbreviations to full forms
```{r}
df <- df %>%
  mutate(
    experience_level = recode(experience_level,
                              "EN" = "Entry", 
                              "MI" = "Mid",
                              "SE" = "Senior", 
                              "EX" = "Executive"
    ),
    
    company_size = recode(company_size, 
                          "S" = "Small", 
                          "M" = "Medium", 
                          "L" = "Large"
    ),
    
    employment_type = recode(employment_type, 
                             "PT" = "Part-time", 
                             "FT" = "Full-time", 
                             "FL" = "Freelance", 
                             "CT" = "Contract"
    )
  )
```

### M2: map all country codes to full names
```{r}
# map country full names
countrycode <- read.csv("data/countrycode.csv")
#head(countrycode)

# inner join
df <- df %>%
  left_join(countrycode, by = c("employee_residence"="code")) %>%
  left_join(countrycode, by = c("company_location"="code")) %>%
  rename(
    country_employee = country.x,   #new = old
    country_company = country.y
  ) %>%
  select(-employee_residence, -company_location)
#head(df, 5)
```


### M3: get job category
```{r}
# read data
job_cat <- read.csv("data/jobtitle_vs_category.csv")


# get unique value
unique_title_vs_category <- job_cat %>%
  distinct(job_title, job_category) %>%   #drop duplicates
  arrange(job_title)

head(unique_title_vs_category, 5)


# left join by job_title
df <- df %>%
  left_join(unique_title_vs_category, by="job_title")
```

```{r}
# check na: 337 cols
colSums(is.na(df))

# check unique values of category
unique_title_vs_category %>%
  distinct(job_category)   # 10 different categories


# manually map "job_title" to "job_category"
df <- df %>%
  mutate(
    job_category = case_when(
      is.na(job_category) & str_detect(job_title, "AI|ML") ~ "Machine Learning and AI",
      is.na(job_category) & str_detect(job_title, "Data Analyst|Analytics") ~ "Data Analysis",
      is.na(job_category) & str_detect(job_title, "Business Intelligence|Data Reporting") ~ "BI and Visualization",
      is.na(job_category) & str_detect(job_title, "Prompt Engineer") ~ "Prompt Engineer",
      is.na(job_category) & str_detect(job_title, "Data Management|Principal") ~ "Data Management and Strategy",
      is.na(job_category) & str_detect(job_title, "Data Science|Research|Scientist") ~ "Data Science and Research",
      is.na(job_category) & str_detect(job_title, "Quality|Operations") ~ "Data Quality and Operations",
      is.na(job_category) & str_detect(job_title, "Pipeline|Robotics|Developer") ~ "Data Engineering",
      is.na(job_category) ~ "Computational Biologist",   # only one
      TRUE ~ job_category
    )
  )
```

### get 
```{r}
# unify country name
library(maps)
library(countrycode)

middle_east <- c("Saudi Arabia", "United Arab Emirates", "Iran", "Iraq", "Israel", "Turkey", "Egypt", "Qatar", "Kuwait")

df <- df %>%
  mutate(country_company = countrycode(country_company, origin = "country.name", destination = "country.name")) %>%
  mutate(continent = countrycode(country_company, origin = "country.name", destination = "continent"))

head(df, 5)
```


```{r}
# final adjustment
df <- df %>%
  select("work_year", "experience_level", "employment_type", "job_category", 
         "job_title", "salary_in_usd", "company_size", "continent", 
         "country_company") %>%
  rename("year" = "work_year",
         "company_country" = "country_company", 
         "job_type" = "employment_type", 
         "salary_usd" = "salary_in_usd")

# save as RData
#save(df, file = "data/data_jobs_salary.RData")
```


```{r}
# test
load("data/data_jobs_salary.RData")
head(df, 109)


# basic info
summary(df)
describe(df)
str(df)


# simple stats of num cols
df %>%
  select("salary_usd") %>%
  describe() %>%
  t() %>%
  as.data.frame()
```


---

## Plot
```{r}
# Salary in USD vs. Job Category (mid level)
df %>%
  filter(experience_level == "Mid") %>%
  ggplot(aes(x=salary_usd, y=job_category)) + 
  geom_boxplot() + 
  scale_x_continuous(labels = scales::comma) +
  labs(
    title = "Salary in USD vs. Job Category (mid level)",
    x = "Salary (in USD)",
    y = "Job Category"
  ) + 
  theme_minimal()
```


```{r}
# Salary in USD vs. Work Year
df %>%
  filter(job_title == "Data Analyst") %>%
  ggplot(aes(x=year, y=salary_usd, fill=experience_level)) + 
  geom_boxplot() + 
  scale_y_continuous(labels = scales::comma) + 
  coord_cartesian(ylim=c(0, 300000)) + 
  labs(
    title = "Data Analyst Salary Trends by Experience Level", 
    x = "Year", 
    y = "Salary in USD"
  )
```


```{r}
# Median salary in usd vs. Country
df %>%
  group_by(company_country) %>%
  summarize(median_country = median(salary_usd)) %>%
  arrange(desc(median_country)) %>% 
  slice_head(n = 15) %>%
  ggplot(aes(x=median_country, y=reorder(company_country, median_country), fill = median_country)) + 
  geom_col() + 
  scale_x_continuous(labels = scales::comma) +
  labs(
    title = "Top 15 Countries by Median Salary in USD",
    x = "Median Salary (in USD)",
    y = "Country"
  ) +
  theme_minimal()
```


```{r}
# Trend of Job Category
g <- df %>%
  group_by(year, job_category) %>%
  #count(job_category, name="job_count")  %>%
  summarise(job_count = n(), name="job_count") %>% 
  ggplot(aes(x=year, y=job_count, color=job_category)) + 
  geom_line() + 
  geom_point() +
  scale_y_log10() +
  labs(
    title = "Trend of Job Category",
    x = "Year",
    y = "Number of Job Categories"
  ) +
  theme_minimal()

ggplotly(g)
```



```{r}
library(maps)
library(countrycode)
library(viridis)
library(scales)


world_map <- map_data("world")

# calculate median
df_map <- df %>%
  group_by(company_country) %>%
  summarise(median_salary = median(salary_usd, na.rm = TRUE)) %>%   #drop NA
  arrange(desc(median_salary))

# unify convert country name in world_map
world_map$region <- countrycode(world_map$region, origin = "country.name", destination = "country.name")
# join location data and salary data
df_map_salary <- left_join(world_map, df_map, by = c("region" = "company_country"))


# plot
p <- ggplot() +
  geom_polygon(data = df_map_salary,
               aes(x = long, y = lat,
                   group = group, fill = median_salary, 
                   text = paste(region, "\nAvg Salary: $", round(median_salary, 2))
               ), 
               color = "white", size = 0.2) +
  scale_fill_viridis_c(option = "magma", na.value = "gray90", labels = comma) +  
  theme_minimal() +
  theme(
    legend.key.size = unit(0.4, "cm"),  
    legend.text = element_text(size = 8),  
    legend.title = element_text(size = 9, face = "bold"), 
    axis.title.x = element_blank(), 
    axis.title.y = element_blank()
  ) +
  labs(title = "Median Salary Distribution (in USD)", fill = "USD")


ggplotly(p, tooltip = "text", dynamicTicks = TRUE)
```

